---
title: "Hands on with Regularization and Tree based Model Ensemble"
output: html_notebook
---


```{r}
housing_df = read.csv("housing.csv")[,-1]
## first column is Id so it can be removed
```

## Section 1. Data Cleaning

__1.__ Take a summary of the data and explore the result. How many categorical and numerical variables are
there in the dataset?

```{r}
summary(housing_df)
```
```{r}
length(housing_df[,sapply(housing_df, is.character) == TRUE]) ## Number of categorical variables
length(housing_df[,sapply(housing_df, is.numeric) == TRUE]) ## Number of numerical variables
```

__2.__ (1pt) Which columns have missing values and what percentage of those columns have NAs? (Note.
You can use colMeans(is.na(your data frame)) to find the percentage of NAs in each column).
```{r}
missing_vals_columns = which(colSums(is.na(housing_df)) > 0)
sort(colSums(sapply(housing_df[missing_vals_columns], is.na)), decreasing = TRUE)
```
The above columns have missing values
```{r}
means = colMeans(sapply(housing_df[missing_vals_columns], is.na))
newmeans = as.numeric(sprintf("%2.3f", means)) * 100
names(newmeans) = names(means)
sort(newmeans, decreasing = TRUE)
```
The above are the % NAs in each column with NAs
